---
title: "Machine learning CA 2"
author: "Namil Muhammed - 10506569"
date: "3 August 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
## Start H2O
library(h2o)
library(ggplot2)

```
```{r}
h2o.init()                  ##Use the h2o.init() method to initialize H2O.

h2o.init(nthreads = -1)   
h2o.clusterInfo()

```
####This below package provides an easy and simple way to read, write and display bitmap images stored in the JPEG format. It can read and write both files and in-memory raw vectors.
```{r}
library(jpeg)
```
####Read your machine's bird image
```{r}
birdimg <- readJPEG("E:/second semester/machine learinig/Bird Image for CA2.jpg")
head(birdimg)              ##first six rows corresponding to the image
```

```{r}
birdimgDM <- dim(birdimg)        ##Check Dimensions ofthe bird Image
birdimgDM

birdimgRGB <- data.frame(                    ##Assign RGB Channels to the birdImage data frame
  x=rep(1:birdimgDM[2], each=birdimgDM[1]),
  y=rep(birdimgDM[1]:1, birdimgDM[2]),
  R= as.vector(birdimg[,,1]),
  G= as.vector(birdimg[,,2]),
  B= as.vector(birdimg[,,3])
)
head(birdimgRGB)                ##first six rows corresponding to the RGB bird image

```
####Using ggplot theme to plot a image
```{r}
plotTheme <- function(){
  theme(
    panel.background=element_rect(size=3,
                                  colour = "black",
                                  fill = "white"),
    axis.ticks=element_line(size = 2),
    panel.grid.major = element_line(colour = "gray80",
                                    linetype = "dotted"),
    panel.grid.minor = element_line(colour = "gray90",
                                    linetype = "dashed"),
    
    axis.title.x = element_text(size = rel(1.2),
                                face = "bold"),
    axis.title.y = element_text(size = rel(1.2),
                                face = "bold"),
    plot.title = element_text(size = 20, face = "bold",
                              vjust = 1.5)
  )
}
##Plotting the RGB Image of Parrot
ggplot(data = birdimgRGB, aes(x=x, y=y))+
  geom_point(colour=rgb(birdimgRGB[c("R", "G", "B")]))+
  labs(title = "Original Image: Bird")+
  xlab("x")+
  ylab("y")+
  plotTheme()

```

```{r}
##Display RGB Image Summary
bird.hex <- as.h2o(birdimgRGB)

head(bird.hex)
```
####Applying k-means clustering to the bird image through h2o
####k means clustering with k=2 
```{r}
kmeans.2 <- h2o.kmeans(training_frame = bird.hex, k = 2, x = c('R','G','B'))

kmeans.2@model$centers          #Dispay data with k=2 clusters

##Predict model using k=2 function with h2o.predict
kmeans2cluster <- h2o.predict(kmeans.2, bird.hex[,c('R','G','B')])

head(as.data.frame(kmeans.2@model$centers[as.vector(kmeans2cluster)+1,]))


```
```{r}
rgbcluster2 <- as.data.frame(kmeans.2@model$centers[as.vector(kmeans2cluster)+1,])[,c('r','g','b')]
kcolors.2 <- rgb(rgbcluster2)

##Plot the image with k=2 clusters

ggplot(data = birdimgRGB, aes(x=x, y=y))+
  geom_point(colour=kcolors.2)+
  labs(title = paste(2, "Colours"))+
  xlab("x")+
  ylab("y")+
  plotTheme()
```

```{r}
##k means image clustering with k=3
kmeans.3 <- h2o.kmeans(training_frame = bird.hex, k = 3, x = c('R','G','B'))

kmeans.3@model$centers           #Display data with k=3 clusters

##Predict model using k=3 function with h2o.predict

kmeans3cluster <- h2o.predict(kmeans.3, bird.hex[,c('R','G','B')])    

head(as.data.frame(kmeans.3@model$centers[as.vector(kmeans3cluster)+1,]))
```
```{r}
rgbcluster3 <- as.data.frame(kmeans.3@model$centers[as.vector(kmeans3cluster)+1,])[,c('r','g','b')]
kcolors.3 <- rgb(rgbcluster3)

##Plot the image with k=3 clusters

ggplot(data = birdimgRGB, aes(x=x, y=y))+
  geom_point(colour=kcolors.3)+
  labs(title = paste(3, "Colours"))+
  xlab("x")+
  ylab("y")+
  plotTheme()
```

```{r}
##k means image clustering with k=4
kmeans.4 <- h2o.kmeans(training_frame = bird.hex, k = 4, x = c('R','G','B'))

kmeans.4@model$centers     ##data with k=4 clusters

##Predict model using k=4 function with h2o.predict

kmeans4cluster <- h2o.predict(kmeans.4, bird.hex[,c('R','G','B')])

head(as.data.frame(kmeans.4@model$centers[as.vector(kmeans4cluster)+1,]))
```
```{r}
rgbcluster4 <- as.data.frame(kmeans.4@model$centers[as.vector(kmeans4cluster)+1,])[,c('r','g','b')]
kcolors.4 <- rgb(rgbcluster4)

##Plot the image with k=4 clusters
ggplot(data = birdimgRGB, aes(x=x, y=y))+
  geom_point(colour=kcolors.4)+
  labs(title = paste(4, "Colours"))+
  xlab("x")+
  ylab("y")+
  plotTheme()
```

```{r}
##k means image clustering with k=5
kmeans.5 <- h2o.kmeans(training_frame = bird.hex, k = 5, x = c('R','G','B'))

kmeans.5@model$centers        ##Display data with k=5 clusters

##Predict model using k=5 function with h2o.predict

kmeans5cluster <- h2o.predict(kmeans.5, bird.hex[,c('R','G','B')])

head(as.data.frame(kmeans.5@model$centers[as.vector(kmeans5cluster)+1,]))
```

```{r}
rgbcluster5 <- as.data.frame(kmeans.5@model$centers[as.vector(kmeans5cluster)+1,])[,c('r','g','b')]
kcolors.5 <- rgb(rgbcluster5)
##Plot the image with k=5 clusters
ggplot(data = birdimgRGB, aes(x=x, y=y))+
  geom_point(colour=kcolors.5)+
  labs(title = paste(5, "Colours"))+
  xlab("x")+
  ylab("y")+
  plotTheme()
```
 
```{r}
## function for Compare predicted image and actual image
sse <- function(predicted, actuals) {
  sum ( (predicted-actuals)^2, na.rm=TRUE ) 
}
```
###Elbow Method for finding optimal value of k
###Creates Elbow for 2,3,4,5 clusters of Image
```{r}
elbow <- c(sse(rgbcluster2, birdimgRGB[,c('R','G','B')]),
           sse(rgbcluster3, birdimgRGB[,c('R','G','B')]),
           sse(rgbcluster4, birdimgRGB[,c('R','G','B')]),
           sse(rgbcluster5, birdimgRGB[,c('R','G','B')]))
##Plot line graph of Elbow using both line and points to see 
##optimal value of k for clusters by using elbow method
plot(elbow, type = 'b')
```

```{r}
##function to compute total-within-cluster
k <- 1:15
elbow <- vector()
for (i in k){
  kmeans <- h2o.kmeans(training_frame = bird.hex, k = i, x = c('R','G','B'))
  kmeanscluster <- h2o.predict(kmeans, bird.hex[,c('R','G','B')])
  rgbcluster <- as.data.frame(kmeans@model$centers[as.vector(kmeanscluster)+1,])[,c('r','g','b')]
  elbow <- append(elbow, sse(rgbcluster, birdimgRGB[,c('R','G','B')]))
}

#Extract elbow for 1-15 clusters(k elbow)

plot(elbow, type='b', xlab = 'Number of Clusters', ylab = 'Sum of Squared Errors')

```

###From this plot we can see that the optimal value is k=4 by using elbow method 
###because after k=4 there is no significant change.